%% -*- fill-column: 72; eval: (auto-fill-mode -1); eval: (visual-fill-column-mode 1); eval: (visual-line-mode 1); eval: (adaptive-wrap-prefix-mode 1) -*-
\documentclass[10pt, a4paper, landscape]{article}
\usepackage[T1]{fontenc}
\usepackage{multicol}
\setlength{\columnsep}{5mm}
\usepackage[medium, compact]{titlesec}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\usepackage{beton}
\DeclareFontSeriesDefault[rm]{bf}{sbc}
\usepackage{eulervm}
\usepackage{amsmath}
\usepackage[margin=0.51in]{geometry}
\usepackage{parskip}
\usepackage{tabularx}
\usepackage{array}
\usepackage{booktabs}
\usepackage{microtype}
%\usepackage{fancyhdr}
%%
%%\usepackage[style=authoryear]{biblatex}
%%\addbibresource{notes.bib}
%%
%%
\newcommand{\defn}[1]{\textbf{#1}}
\newcommand{\isdef}{\stackrel{\text{def}}{=}}
\newcommand{\set}[1]{\mathbold{#1}}
\newcommand{\imag}{\mathrm{i}}
%%
\title{All the maths we know}
\author{}
\date{21 October 2023}
%%
\renewcommand{\arraystretch}{1.2}
\setlength{\tabcolsep}{3pt}
%%
\begin{document}%\maketitle
%%
\begin{multicols*}{3}\raggedcolumns%
%%
\maketitle

\section*{Set theory}
\subsection*{Sets}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule
  \defn{Set} & A collection of things, called its \emph{elements} (or \emph{members}). \\

  $x \in X$    & ``$x$ is an element of set $X$.'' \\

  $X\cup Y$     & The \defn{union} of $X$ and $Y$: all elements of $X$ together with all elements of $Y$. \\

  $X\cap Y$     & The \defn{intersection} of $X$ and $Y$; its elements are
  those that are elements of both $X$ and $Y$. \\

  $X \setminus Y$      & The \defn{set difference}: all elements of $X$ that are not elements of $Y$. Sometimes $X-Y$. \\

  $X \subset Y$    & ``Every element of $X$ is an element of $Y$.'' $X$ is a \emph{subset} of~$Y$. \\

  $\{x\in X\mid P(x)\}$ & The set of those elements of $X$ satisfying the
  predicate $P$ (\textit{i.e.}, for which $P(x)$ is true). \\

  $\emptyset$        & The \defn{empty set}---the set with no members. \\

\end{tabularx}

%% ------------------------------------------------------------

\subsection*{Pairs and tuples}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule
  \defn{Pair} & An ordered list of two things: like a set but the order matters and they can be the same thing. For example: $(1, 2)$. \\
  
  \defn{$n$-tuple} & An ordered list of $n$ things. For example: $(x_1, x_2,\dotsc, x_n)$. \\

  $X\times Y$      & The \defn{Cartesian product} of $X$ and $Y$ is the set
  of all pairs $(x,y)$ where $x \in X$ and $y \in Y$. \\

  $X^n$ & The set of all $n$-tuples of elements of~$X$. The Cartesian product of $X$ with itself $n$ times. \\

\end{tabularx}

%% ------------------------------------------------------------

\subsection*{Maps}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule
  \defn{Map} & A rule which assigns, to every element of a set (called the \emph{domain}), an element of another set (called the \emph{codomain} or \emph{range}). Sometimes called a \emph{function}, especially when the codomain is numbers. \\

  $f\colon X\to Y$   & ``$f$ is a map from $X$ to $Y$.'' \\

  $f\colon x \mapsto y$  & ``Specifically, $f$ maps the element $x \in X$ to the
  element $y \in Y$.'' \\
  $f \circ g$    & The map $g$ followed by the map $f$. \\

  \defn{Injection} & A map, $f\colon X \to Y$, is \emph{injective} (also
  \emph{one-to-one}) if no more than one element of $X$ maps to a
  particular element of $Y$. \\

  \defn{Surjection} & A map, $f\colon X \to Y$, is \emph{surjective} (also \emph{onto}) if every element of $Y$ is mapped to by some element of $X$. \\

  \defn{Bijection} & A map, $f\colon X \to Y$, is \emph{bijective} if it
  is both injective and surjective (also “one-to-one and onto”). \\

  $X\cong Y$ & There exists a bijection between sets $X$ and~$Y$. Equivantly: $X$ and $Y$ are \emph{isomorphic as sets}. 
\end{tabularx}


%% ============================================================

\section*{Useful maps}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule
  \defn{Operator} & A \emph{binary operator} on a set $\set{X}$ is a map $\star \colon \set{X}\times\set{X} \to \set{X}$. That is, a binary operator takes two elements of $\set{X}$ and returns an element of~$\set{X}$. Operators are typically written in “infix” notation, $x\star y$, rather than in a function notation, $\star(x,y)$. \\

  \defn{Associative} & A binary operator, $\star$, is \emph{associative} if
  \begin{equation*}
    (a \star b) \star c = a \star (b \star c).
  \end{equation*}
  Almost all operators you will meet are associative. Since the order in which the operations are carried out doesn't affect the result, the parentheses are often omitted, as in $a\star b\star c$.\\
  
  \defn{Commutative} & A binary operator, $\star$, is \emph{commutative} if
  \begin{equation*}
    a \star b = b \star a.
  \end{equation*}
\end{tabularx}

%% ============================================================

\section*{Numbers}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule
  $\set{N}$ & The \emph{natural numbers:} $0, 1, 2, \dotsc$. \\
  $\set{Z}$ & The \emph{integers:} $\dotsc, -2, -1, 0, 1, 2, \dotsc$. \\
  $\set{Q}$ & The \emph{rationals:} All numbers that can be written
  as $m/n$ where $m$ and $n$ are integers. \\
  $\set{R}$ & The \emph{reals:} The rationals and ``all the numbers in
  between.'' \\
  $\set{R}^+$ & The non-negative reals. \\
  $\set{C}$ & The \emph{complex numbers:} Numbers of the form $a + b\imag$, where $a$ and $b$ are real numbers and $\imag^2=-1$. 
\end{tabularx}

%% ============================================================

\section*{Vector spaces}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule \defn{Vector space} & A \emph{real vector space} is a set,
  $V$, together with: (i) a commutative, associative, binary operator,
  $+$, on $V$; (ii) a map, $\cdot\colon \set{R}\times V\to V$; and (iii) a
  distinguished element $\mathbold{0}\in V$, such that:
  \begin{enumerate}
  \item $v + \mathbold{0} = v$ for all $v\in V$;
  \item For any $v\in V$ there exists an element $-v \in V$ such that $v+(-v)=\mathbold{0}$;
  \item $1 \cdot v = v$ for all $v\in V$;
  \item $\alpha\cdot(\beta\cdot v) = (\alpha\beta)\cdot v$ for all $\alpha,\beta\in \set{R}$ and all $v\in V$.
  \item $\alpha\cdot(v+w) = \alpha\cdot v+ \alpha\cdot w$ and $(\alpha+\beta)\cdot v = \alpha\cdot v + \beta \cdot v$ for all $\alpha, \beta\in \set{R}$ and $v,w \in V$.
  \end{enumerate}

  Replacing $\set{R}$ with $\set{C}$ in the above we obtain a
  \emph{complex vector space}. \\

  \defn{Vector} & A vector is an element of a vector space.
\end{tabularx}

\section*{Examples of vector spaces}
\begin{tabularx}{\columnwidth}{@{}l>{\raggedright\arraybackslash}X@{}}
  \toprule
  $\set{R}^n$ (or $\set{C}^n$) & The set of $n$-tuples of $\set{R}$ (or $\set{C}$), together with the operation of “element-wise” addition:
  \begin{multline*}
    (x_1, \dotsc, x_n) + (y_1, \dotsc, y_n) \isdef \\ (x_1+y_1, \dotsc, x_n+y_n) 
  \end{multline*}
  and multiplication by $\set{R}$ (or $\set{C}$):
  \begin{equation*}
    \lambda \cdot (x_1, \dotsc, x_n) \isdef (\lambda x_1, \dotsc, \lambda x_n) 
  \end{equation*} \\

  $\set{R}^X$ (or $\set{C}^X$) & For $X$ a set, $\set{R}^X$ denotes the set of all functions $X\to\mathbold{R}$, together with the operation of “pointwise addition”:
  \begin{equation*}
    (f+g)(x) \isdef f(x) + g(x)
  \end{equation*}
  and “pointwise multiplication by $\mathbold{R}$”:
  \begin{equation*}
    (\alpha f)(x) \isdef \alpha f(x).
  \end{equation*}
  (The notation “$(f+g)(x)$” means “the function $f+g$, where $+$ is addition of functions, evaluated at the point $x$.”)
    
\end{tabularx}

\end{multicols*}
\end{document}
